---
title: "Raster Data"
subtitle: "ESM 244 - Assignment 3 Task 2"
author: "Minnie Ringland"
date: "Feb 25, 2021"
output:
  html_document:
    theme: yeti
    code_folding: hide
    highlight: haddock
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(raster) ### NOTE: attaching this BEFORE tidyverse
library(tidyverse)
library(here)
library(sf)
library(fasterize)
library(rnaturalearth)
#library(rgeos)
```
***  


### About the Data

AquaMaps (Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander,
S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic
species. www.aquamaps.org), showing the probability of occurrence (based on relative
environmental suitability, including species preferences for water temperature, depth,
salinity, and distance to land) of 35 cetacean species that can be found in the California
Bight. The extent of the rasters is 125° W to 115°W (which R considers -125 to -115)
and 32°N to 38°N.


Create a raster of species richness of cetacean species off the coast of
California. Species richness is simply the number of species that occur in a
particular location.
○ With probability of occurrence data such as AquaMaps, it is common to
decide on a threshold value above which a species is considered
“present.” Choose a probability threshold (a threshold of .6, or higher will
result in more interesting maps), and use your threshold to reclassify each
probability layer to a “presence” layer of either 0 (not present) or 1
(present).

```{r, cache = TRUE}
# Create a list of files to use and rasterize as a stack

aquamap_files <- list.files(path = here("ca_cetaceans"), full.names = TRUE, pattern = "*.tif")

aq_stack <- raster::stack(aquamap_files)
smaller_aq <- raster::aggregate(aq_stack, fact = 2, fun = mean) # Use this smaller/faster version to build code

plot(smaller_aq)
```

Set a threshold for presence probability --> 0.7:

```{r set presence threshold}

is_present <- function(x, thresh = 0.7) {
  y <- ifelse(x >= thresh, 1, 0)
  return(y)
}

presence <- calc(smaller_aq, fun = is_present)
plot(presence, col = 'blue4')

```

***
```{r coastline}
coast <- ne_coastline(returnclass = "sf")

coast_df <- coast %>% 
  as.data.frame()

#coast_df <- raster::rasterToPoints(coast) %>%
#  as.data.frame()

```


`ggplot` likes dataframes - not rasters - so we have to convert our raster to a dataframe. 
```{r calculate richness}

richness <- calc(presence, fun = sum, na.rm = TRUE)

species_df <- raster::rasterToPoints(richness) %>%
  as.data.frame()

```



```{r plotting, fig.cap="**Figure 1.** Cetacean Species Richness along the California Coast, using threshold probability > 0.7"}

ggplot() +
  geom_raster(data = species_df, aes(x = x, y = y, fill = layer), alpha = 0.6) +
  geom_sf(data = coast, size = 1.5, color = "grey", label = "California Coastline") +
  coord_sf(xlim = c(-130, -114), ylim = c(31, 39)) +
  scale_fill_gradient(low = 'lemonchiffon1', high = 'lawngreen') +
  labs(x = "Longitude", y = "Latitude", title = "Cetacean Species Richness", fill = "Number of Species Present") +
  theme_minimal() #+
  #theme(panel.background = element_rect(fill = 'slateblue4'))

```

min(species_df$x), max(species_df$x)),
           ylim = c(min(species_df$y), max(species_df$y))

***

**Data Source:**

