---
title: "Text Sentiment Analysis of the Resilient Merced Report"
subtitle: "ESM 244 - Assignment 3 Task 1"
author: "Minnie Ringland"
date: "Feb 25, 2021"
output:
  html_document:
    theme: paper
    code_folding: hide
    highlight: espresso
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse) #for everything
library(here) # for file reference
library(tidytext) # for text analysis
#library(textdata)
library(pdftools) # for reading in PDFs
library(kableExtra) # for making tables
library(ggwordcloud) # for wordclouds

```

***  


### About the Report

In 2016, two California senate bills were passed that identified conservation and management of **natural and working lands** as a key strategy to meeting emission goals. In response, the California Air Resources Board (CARB) [https://ww2.arb.ca.gov/nwl-inventory], which spearheads the state’s climate response, laid out goals to develop a statewide carbon inventory. Later that year, the County of Merced partnered with the state Department of Conservation and The Nature Conservancy to pilot a scenario planning tool to allow individual jurisdictions to quantify and project carbon storage in natural and working lands on a more local level.

This tool is called TerraCount, and the Resilient Merced report describes how the tool was developed, how it can be used, and the results of the study in Merced County.

Today, we'll be exploring the text of the report to highlight key terms, and analyze the tone or sentiment of the report.


```{r read PDFs, cache = TRUE}
# Read in the pdfs
merced_text <- pdf_text(here("pdfs","resilientcountiesguide_1_.pdf"))
sb_text <- pdf_text(here("pdfs","2016_ghg.pdf"))
```


***


The PDF text is initially read in as a character vector - a single, very long row of text. To wrangle this into a workable dataset, I need to:  
- convert to tidyverse dataframe  
- use `str_split()` to break by line (using '\n')  
- remove excess white space  
- remove introductory text (table of contents, acknowledgments, etc)  
- split by section/chapter  
- split each word into its own row  
- remove stopwords  


```{r tidy and wrangle}
# Turn into a tidyverse df, split by line
merced_tidy <- data.frame(merced_text) %>% 
  mutate(text_full = str_split(merced_text, pattern = '\\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full)) %>% # remove white space
  slice(-(1:188)) %>% # slice to remove doc intro
  mutate(section = str_extract(
                      case_when(
                      str_detect(text_full, pattern = "Section \\d:") ~ text_full, # split by Section (chapter)
                      TRUE ~ NA_character_),
                      "\\d")) %>% 
  fill(section)

#merced_test <- head(merced_tidy,1)
#pattern <- "Section \\d:"
#str_detect(merced_test, pattern)
#unique(merced_tidy$section)

# Finally, split so that each word has it's own row, i.e. "tokenize"
merced_tokens <- merced_tidy %>% 
  unnest_tokens(word, text_full) %>% 
  select(-merced_text) %>% 
  anti_join(stop_words) # and remove stopwords

```

We are left with a dataset containing only the words of interest, organized by document section. We can see that each section use the following number of unique words.

```{r unique words}
merced_wordcount <- merced_tokens %>% 
  group_by(section) %>% 
  summarize(words = n_distinct(word))

kable(merced_wordcount, caption = "Table 1. Unique words in each section of Resilient Merced") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))
```

***  
### Most Frequently Used Words

```{r top 5 and top 50}
merced_counts <- merced_tokens %>% 
  count(section, word)

top_5_words <- merced_counts %>% 
  group_by(section) %>% 
  arrange(-n) %>% 
  slice(1:5)

ggplot(data = top_5_words, aes(x= word, y = n)) +
  geom_col(fill = "blue") +
  facet_wrap(~section, scales = "free") +
  coord_flip()

merced_top50 <- merced_counts %>% 
  arrange(-n) %>% 
  slice(1:50)

ggplot(data = merced_top50, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "square") +
  scale_size_area() +
  scale_color_gradientn(colors = c("darkgreen","green")) +
  theme_minimal()

```



***  
### Sentiment Analysis

```{r sentiment analysis with afinn}
merced_afinn <- merced_tokens %>% 
  inner_join(get_sentiments("afinn"))

afinn_counts <- merced_afinn %>% 
  count(section, value)

ggplot(data = afinn_counts, aes(x = value, y = n)) +
  geom_col() +
  facet_wrap(~section)

# Find the mean afinn score by chapter: 
afinn_means <- merced_afinn %>% 
  group_by(section) %>% 
  summarize(mean_afinn = mean(value))

afinn_plot <- ggplot(data = afinn_means,
       aes(x = section, y = mean_afinn)) +
  geom_col() +
  coord_flip()

```

```{r sentiment analysis with bing}
merced_bing <- merced_tokens %>% 
  inner_join(get_sentiments("bing"))

bing_counts <- merced_bing %>% 
  count(section, sentiment)

ggplot(data = bing_counts, aes(x = sentiment, y = n)) +
  geom_col() +
  facet_wrap(~section)

```
```{r sentiment analysis with nrc}
merced_nrc <- merced_tokens %>% 
  inner_join(get_sentiments("nrc"))

nrc_counts <- merced_nrc %>% 
  count(section, sentiment)

ggplot(data = nrc_counts, aes(x = sentiment, y = n)) +
  geom_col() +
  facet_wrap(~section) +
  coord_flip()
```

***

**Data Source:**

*Resilient Merced: A County Guide to Advance Climate Change Mitigation and Complementary Benefits through Land Management and Conservation.* California Department of Conservation and The Nature Conservancy. Published 2019. Available for download at https://maps.conservation.ca.gov/TerraCount/downloads/